---
title: "Collecting Census Data"
author: "CKelling"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will start by loading the tidycensus library, which has functions that will allow us to download Census data.
```{r, message = FALSE}
# library for downloading Census variables
library(tidycensus)

#for formatting data
library(tidyverse)
library(maditr)
```

Next, you will need to sign up for a Census API key. Please read the terms of service! For the organization, you can put Carleton College. Also be sure to never share this key- it is a unique identifier to you! You can paste it below, run the function, and then delete if from your code.

To sign up for a Census API key go to the following website: \url{https://api.census.gov/data/key_signup.html}.


After you sign up, we will install your census API key into your R environment. You can run this once and then delete your API key and comment the rest of the code out.
```{r}
#census_api_key("", install = TRUE)

#check to make sure your census key is there:
#Sys.getenv("CENSUS_API_KEY")
```

Now, you should choose the list of variables that you want to download. I will include two variables here, so you can see how this works. The full list of variables is included here: \url{https://api.census.gov/data/2017/acs/acs5/variables.html}. 

You should copy the code for the variable into the following function.

Below, I've included information on how to get county-level Census data for two variables from Ohio for the year 2018 (the most recent ACS to this date- the dates will be included in the output). The "E" indicates that these are the point estimates (not uncertainty estimates).
```{r}
#download ACS (Census) data
oh_census <- get_acs(geography = "county", 
                       variables = c("B19001_001E", 
                                     "B01003_001E"), 
                     year = 2018,
                     state = "OH",
                     geometry = F)
```


Now, we need to create a codebook for the data so we can keep track of the variable names- the Census codes don't tell us anything! We will join it to the census data and then make it into a format where rows are counties and columns are census variables.
```{r}
#creating a codebook for you to  track of the variables.
code_book <- rbind(c("B19001_001", "household_income"),
                   c("B01003_001", "total_pop_acs"))

#join to the Census data
code_book <- as.data.frame(code_book)
colnames(code_book) <- c("variable", "var_name")
oh_census <- left_join(oh_census, code_book)

#format the data so there is a row for each census tract and column for every variable
ohio_acs_data <- maditr::dcast(oh_census, NAME ~ var_name, 
                               value.var = "estimate", 
                               fun.aggregate = NULL)

#take a look at the dataset
head(ohio_acs_data)
```

Next, you could join the Ohio ACS data that we just created with your incarceration dataset. Take a look at the coding tips on the assignment pdf!

```{r}
incarc_data <- read.csv("https://github.com/vera-institute/incarceration-trends/raw/master/incarceration_trends.csv")
tx_incarc <- incarc_data %>%
  filter(state == "TX")
```

